Semi-Permeable Vesicles: Mathematical Formulation, Numerics, and Physics

### 2020-08-31 vesicle_code/docs/permeableVes/main.tex
updating pnas paper

\section*{Governing equations} \label{sc:formulate}
Consider a two-dimensional semi-permeable vesicle suspended in an
unbounded viscous fluid domain, subjected to an imposed flow
$\uu_\infty(\xx)$, for any $\xx \in \RR^2$.  Assume that the interior
and exterior fluids have the same viscosity $\mu$. Let $\xx$ be the
position of the interface $\gamma$, $\uu$ be the fluid velocity, and
$p$ be the pressure. In the vanishing Reynolds number limit, the
governing equations for the ambient fluid are
%
\begin{alignat}{3}
  \label{eqn:governing}
  -\nabla p + \mu \triangle \uu &= 0, 
  \nabla \cdot \uu = 0, \qquad &&\xx \in \RR^2 \setminus \gamma, \\
%
  \uu(\xx) &\rightarrow \uu_\infty(\xx),
    &&\norm{\xx} \rightarrow \infty,
\end{alignat}
where $\mu$ is the fluid viscosity. Letting $\nn$ be the outward normal
to $\gamma$, $\jump{\cdot}$ be the jump across the interface, and $T$ be
the hydrodynamic stress tensor, the interfacial boundary conditions are
\begin{align}
  \label{eqn:jump}
  \jump{T\cdot\nn} = \ff_\mathrm{mem}, \quad \jump{\uu} = 0,
\end{align}
where $\ff_\mathrm{mem}$ is the membrane force that results from the
Helfrich energy with bending modulus $k_b$, and a tension term that acts
as a Lagrange multiplier to satisfy the local inextensibility condition.
Following~\cite{vee-gue-zor-bir2009}, we modify the bending tension and
forces as
$\ff_\mathrm{mem} = \ff_\mathrm{ben} +
\ff_\mathrm{ten}$, where
\begin{align}
  \label{eqn:membraneForces}
  \ff_\mathrm{ben} = -k_b \xx_{ssss}, \qquad
  \ff_\mathrm{ten} = (\sigma \xx_s)_s.
\end{align}
The physical tension satisfies $\Lambda = \sigma + 1.5 \kappa^2$, where
$\kappa$ is the membrane curvature.

The kinematic condition is modified to account for flow through the
vesicle. A standard choice for this flux is the difference in mechanical
force~\cite{yao-mor2017}
\begin{align}
  \label{eqn:vesicleFlux}
  \uu - \dot{\xx} = - \beta (\ff_\mathrm{mem} \cdot \nn) \nn, \qquad
  \xx \in \gamma,
\end{align}
where $\beta$ is a semi-permeability rate.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{Integral equation formulation}
Since the flow is governed by the linear Stokes equations, it can be
represented in terms of layer potentials. Given the interfacial boundary
conditions~\eqref{eqn:jump}, the fluid velocity is
\begin{align}
  \uu(\xx) = \uu_\infty(\xx) + \SS[\ff_\mathrm{mem}](\xx), \quad
    \xx \in \Omega,
\end{align}
where $\SS$ is the single-layer potential
\begin{align}
  \SS[\ff](\xx) = \frac{1}{4\pi\mu} \int_{\gamma} \left(
    -\mathbf{I} \log\rho + \frac{\rr \otimes \rr}{\rho^2} \right)
    \ff(\yy) ds_{\yy},
\end{align}
with $\rr = \xx - \yy$ and $\rho = \norm{\rr}$. Imposing the flux
condition~\eqref{eqn:vesicleFlux} and the inextensibility condition, the
vesicle velocity satisfies
\begin{align}
  \label{eqn:vesVelocity}
  \dot{\xx} &= \uu_\infty(\xx) + \beta (\ff_\mathrm{mem}\cdot\nn)\nn
  + \SS [\ff_\mathrm{mem}](\xx),  \quad
  \xx_s \cdot \dot{\xx}_s = 0.
\end{align}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{Dimensionless equations}
The inextensibility condition guarantees that the vesicle's length, $L$,
is conserved, but its area, $A$, is not because of semi-permeability.
Therefore, the reduced area $\nu = 4\pi A/L^2 \in (0,1]$ is not
conserved. We non-dimensionalize~\eqref{eqn:vesVelocity} by introducing
the length scale $R_0 = L/2\pi$ and defining the time scale $\tau = \mu
R_0^3/k_b$, the tension scale $Q = k_b/R_0^2$, and the permeability
scale $W = R_0/\mu$. The background velocity $\uu_\infty$ is
non-dimensionalized with $R_0/\tau$. For all simulations, the bending
modulus is $10^{-19}J$. Using these scales, all equations are written in
dimensionless form.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{Area dynamics}
Following the analysis of Veerapaneni et al.~\cite{vee-raj-bir-pur2009},
we compute the steady-state shape of a semi-permeable vesicle in a
quiescent flow. The steady-state shape is a stationary point of the
Lagrangian
\begin{align}
  \mathcal{L} = \frac{1}{2}\int_{\gamma} \kappa^2 \, ds +
    \Lambda \left(\int_{\gamma} ds  - L \right),
\end{align}
where the Lagrange multiplier $\Lambda$ is the tension. Note that there
is no pressure Lagrange multiplier since the vesicle's area fluctuates.
Taking the variation of $\mathcal{L}$ with respect to $\gamma$ and
setting it to zero, the steady state shape satisfies $\kappa_{ss} +
\frac{1}{2}\kappa^3 - \Lambda \kappa = 0$, with initial conditions
$\kappa(0) = \kappa_0$ and $\kappa_s(0) = 0$.  Integrating once, we have
\begin{align}
  \frac{\kappa_s^2}{2} + \frac{\kappa^4}{8} - 
    \frac{\Lambda}{2}\kappa^2 = \frac{\kappa_0^4}{8} - 
    \frac{\Lambda}{2}\kappa_0^2.
  \label{eqn:curvODE}
\end{align}
Assuming $\kappa_s \neq 0$, this differential equation is separable, and
can be solved analytically in terms of the {\em EllipticF} function
using Mathematica. Unlike in the impermeable
case~\cite{vee-raj-bir-pur2009}, the solution of~\eqref{eqn:curvODE} can
not be made periodic unless $\kappa(s) = \kappa_0$. Therefore, the
steady state shape of a semi-permeable vesicle in a quiescent flow is a
circle. By the inextensibility condition, the steady state vesicle has a
radius of $L/2\pi$, so its curvature and tension are
\begin{align}
  \kappa = \frac{2\pi}{L}, \: \Lambda = \frac{2\pi^2}{L^2}.
  \label{eqn:SSshape}
\end{align}
Note that in contrast to a impermeable vesicle, the final tension is
positive rather than negative.

In addition to computing the steady shape, we also compute the transient
area of the vesicle. Since the area of the vesicle is
\begin{align}
  A(t) = \frac{1}{2}\int_{\gamma} (\xx \cdot \nn)\, ds,
  \label{eqn:area}
\end{align}
its derivative includes terms due to the time derivative of $\xx \cdot
\nn$ and a term due to interface stretching~\cite{lai-tse-hua2008}, but
the stretching term is zero because of the local inextensibility
condition. Therefore,
\begin{align}
  \dot{A}(t) =
  \frac{1}{2} \int_{\gamma} (\dot{\xx} \cdot \nn)\, ds  + 
  \frac{1}{2} \int_{\gamma} (\xx \cdot \dot{\nn})\, ds.
\end{align}
Applying integration by parts to the second integral, it is equivalent
to the first.
%\begin{align}
%  \int_\gamma (\xx \cdot \dot{\nn})\, ds 
%    = \int_\gamma \left(\xx \cdot \dot{\xx}_s^\perp \right)
%    = -\int_\gamma \left(\xx^\perp \cdot \dot{\xx}_s\right)\, ds 
%    = \int_\gamma \left(\xx_s^\perp \cdot \dot{\xx}\right)\, ds 
%    = \int_\gamma (\dot{\xx} \cdot \nn)\, ds.
%\end{align}
Therefore,
\begin{align}
  \dot{A}(t) = \int_\gamma \left(\beta (\ff \cdot \nn)\nn 
    + \SS[\ff]\right) \cdot \nn\, ds 
  = \beta \int_\gamma (\ff \cdot \nn)\, ds,
\end{align}
where the last equality follows from the incompressibility of the
single-layer potential.  Using integration by parts and the
Frenet-Serret identities, we have
\begin{align}
  \dot{A}(t) &= \beta \int_\gamma \left( -\xx_{ssss} + 
    (\sigma \xx_s)_s \right) \cdot \nn \, ds \\
  &= -\beta \int_\gamma \left(-\xx_{sss} + \sigma \xx_s 
    \right) \cdot \kappa \xx_s \, ds \\
  &= -\beta \int_\gamma \left(\xx_{ss} \cdot 
    (\kappa \xx_s)_s + \kappa \sigma \right) \, ds \\
  &= -\beta \int_\gamma \left(\kappa (\xx_{ss} \cdot \xx_{ss}) + 
    \kappa_s (\xx_{ss} \cdot \xx_s) + \kappa \sigma \right) 
    \, ds \\
  &= -\beta \int_\gamma \left(\kappa^3 + \kappa \sigma \right) 
    \, ds = \beta \int_\gamma \left(
    \frac{\kappa^3}{2} - \kappa \Lambda \right) \, ds.
  \label{eqn:areaROC}
\end{align}
This is consistent with the steady state curvature and tension
in~\eqref{eqn:SSshape} since these constant values result in the
integrand in~\eqref{eqn:areaROC} being zero.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{Discretization in Space}
Using an integral equation formulation, the unknowns are the vesicle's
position and tension that we track at a set of $N$ discretization
points. All derivatives are computed with Fourier differentiation, and
the single-layer potential operator is discretized with eighth-order
quadrature~\cite{alp1999}. For constricted flows, nearly-singular
integrals are accurately computed with an interpolation-based quadrature
rule~\cite{qua-bir2014}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{Discretization in Time}
To eliminate stringent time step restrictions, we modify a time stepping
method that discretizes high-order derivatives
semi-implicitly~\cite{vee-gue-zor-bir2009}. Introducing notation for the
differential and projection operators
\begin{alignat}{3}
  &B[\xx]\ff = -\frac{d^4}{ds^4} \ff,  \qquad
  &&T[\xx]\sigma = (\sigma \xx_s)_s, \\
  &D[\xx]\ff = \xx_s \cdot \ff_s, 
  &&P[\xx]\ff = (\ff \cdot \nn) \nn,
\end{alignat}
the dynamics of the vesicle are governed by
\begin{align}
  \dot{\xx} = \uu_{\infty}(\xx) + 
  \beta P(B\xx + T\sigma) + \SS(B\xx + T\sigma), \quad
  D \dot{\xx} = 0.
\end{align}
Letting $B^N$ be the bending operator due to the vesicle configuration
at time $t^N$, and similar notation for the other operators, we use the
first-order semi-implicit time stepping method
\begin{alignat}{3}  
  \frac{\xx^{N+1} - \xx^N}{\Delta t} &= \uu_\infty(\xx^N) 
  &&+ \beta P^N\left(B^N\xx^{N+1} + T^N\sigma^{N+1}\right) \nonumber \\
  & &&+ \SS^N\left(B^N\xx^{N+1} + T^N\sigma^{N+1}\right), \\
  D^N\xx^{N+1} &= 1.
\end{alignat}

To achieve high-order accuracy over long time horizons, we modify an
adaptive spectral deferred correction time stepping
method~\cite{qua-bir2016}. Since the area of the vesicle is not
conserved, we only use the length of the vesicle to estimate the local
truncation error that determines if a time step is accepted or not and
the subsequent time step size.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Examples}
We consider a single semi-permeable vesicle in quiescent, shear, and
parabolic flows. We also consider a semi-permeable vesicle passing
through a constriction. For each example, all vesicles have a length
$L$, but the area is not conserved because of semi-permeability, but it
is bounded by $L^2/4\pi$ which is the area of a circle with
circumference $L$. For each simulation, we apply an SDC second-order
adaptive time stepping method~\cite{qua-bir2016}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{Quiescent Flow} 
We consider a semi-permeable elliptical vesicle with initial reduced
area $\nu = 0.65$ and several different permeability rates in a
quiescent flow. The steady-state of an impermeable vesicle has constant
negative tension~\cite{kra-win-sei-lip1996}, but based
on~\eqref{eqn:SSshape}, a semi-permeable vesicle relaxes to a circle
with positive constant tension $\Lambda = 2\pi^2/L^2$. Since the
governing equation for the area in~\eqref{eqn:areaROC} is proportional
to $\beta$, we expect both the time for the vesicle to deviate from its
initial area, and the time to reach its steady state to be proportional
to $\beta^{-1}$. In Table~\ref{tbl:ellipseRelaxTimes}, we report the
time, $t_\mathrm{steady}$, required for the vesicle's area to be within
1.3\% of its maximum area, and the time, $t_\mathrm{inflate}$, required
for the vesicle's area to exceed 1.3\% of its initial area. Both times
are proportional to $\beta^{-1}$. Next we plot the area of a vesicle
with six different permeability rates in Figure~\ref{fig:ellipseArea}.
The area is computed directly from the vesicle shape (solid lines) and
by integrating~\eqref{eqn:areaROC} (marks) with a first-order method. We
include black dashed lines indicating the times, $t_{\mathrm{steady}}$,
that the vesicle is near its steady state circular shape.
\begin{table}[htp]
  \centering
  \caption{\label{tbl:ellipseRelaxTimes} The time for a semi-permeable
  vesicle to its steady state ($t_\mathrm{steady}$) and to begin
  inflating ($t_\mathrm{inflate}$).}
  \begin{tabular}{C{0.8cm}C{2cm}C{0.8cm}C{2cm}C{0.8cm}}
    $\beta$ & $t_\mathrm{steady}$ & ratio & $t_\mathrm{inflate}$ & ratio \\
    \midrule
    $10^{0}$  & $6.45 \times 10^{0}$ & ---  
              & $2.17 \times 10^{-2}$ & --- \\ 
    $10^{-1}$ & $2.00 \times 10^{1}$ & 3.10 & 
                $1.25 \times 10^{-1}$ & 5.76 \\
    $10^{-2}$ & $9.61 \times 10^{1}$ & 4.81 & 
                $1.59 \times 10^{0}$ & 12.72 \\
    $10^{-3}$ & $8.36 \times 10^{2}$ & 8.71 & 
                $2.19 \times 10^{1}$ & 13.77 \\
    $10^{-4}$ & $8.15 \times 10^{3}$ & 9.74 & 
                $2.14 \times 10^{2}$ & 9.77 \\
    $10^{-5}$ & $8.03 \times 10^{4}$ & 9.85 & 
                $2.20 \times 10^{3}$ & 10.14 \\
    \bottomrule
  \end{tabular}
\end{table}

\begin{figure}[htp]
  \centering
%  \ifTikz
%  \input{figures/ellipseArea.tikz}
%  \fi
  \includegraphics[width=0.8\linewidth]{figures/ellipseArea}
  \caption{\label{fig:ellipseArea} The area of an elliptical
  semi-permeable vesicle, with permeability rate $\beta$ in a quiescent
  flow. The solid lines compute the area using~\eqref{eqn:area}, while
  the marks are computed by integrating~\eqref{eqn:areaROC}. The dashed
  black lines are the times when the vesicle's area is within 1.3\% of
  its maximum value of $4\pi$.}
\end{figure}

We further analyze the transient dynamics for small $\beta$ by
considering the vesicle velocity~\eqref{eqn:vesVelocity} due to
semi-permeability, $\beta(\ff \cdot \nn)\nn$, and force balance,
$\SS[\ff]$. Initially, since $\beta$ is small, the semi-permeability
velocity is much smaller than the force balance velocity and the
velocity resembles an impermeable vesicle. As the
vesicle approaches the steady state shape of an impermeable vesicle, the
force balance velocity decreases below the semi-permeability velocity,
and then the semi-permeable velocity begins to dominate, causing the
vesicle to inflate. As the vesicle becomes circular, the
membrane force and the corresponding velocity both converge to
zero. In Figure~\ref{fig:vesVelocity} we plot the force balance velocity
(blue) and the semi-permeability velocity (red) at three times with
$\beta=10^{-3}$. We also plot the velocity of an identically initialized
impermeable vesicle (green). At $t=0.5$, the semi-permeability velocity is
small relative to the force balance velocity, and the semi-permeable
velocity matches with the impermeable vesicle velocity. At time $t=5$,
the magnitudes of the velocity of the impermeable vesicle and the force
balance velocity are comparable to the magnitude of the
semi-permeable velocity, and the semi-permeable vesicle begins to
inflate. Then, at time $t=50$, the impermeable vesicle is at its steady
state, while the semi-permeable vesicle is still tending to a circular
vesicle. The bottom right plot of Figure~\ref{fig:vesVeocity} is the
maximum norm of the different velocity contributions as a function of
time, and the black vertical lines correspond to the times reported in
other three subplots.

>>>>

\section*{Governing equations} \label{sc:formulate}
We consider a two-dimensional semi-permeable vesicle suspended in an
unbounded viscous fluid driven by the imposed flow $\uu_\infty(\xx)$.
Since fluid is transported across the semi-permeable vesicle membrane,
we only consider fluids with uniform viscosity. Letting $\gamma$ be the
vesicle interface, the fluid velocity $\uu$ and pressure $p$ satisfy
%
\begin{alignat}{3}
  \label{eqn:governing}
  -\nabla p + \mu \triangle \uu &= 0, \quad
  \nabla \cdot \uu = 0, \qquad &&\xx \in \RR^2 \setminus \gamma, \\
%
  \uu(\xx) &\rightarrow \uu_\infty(\xx),
    &&\norm{\xx} \rightarrow \infty,
\end{alignat}
where $\mu$ is the fluid viscosity. Letting $\nn$ be the outward normal
to $\gamma$, $\jump{\cdot}$ be the jump across the interface, and $T$ be
the hydrodynamic stress tensor, the interfacial boundary conditions are
\begin{align}
  \label{eqn:jump}
  \jump{T\cdot\nn} = \ff_\mathrm{mem}, \quad \jump{\uu} = 0,
\end{align}
where $\ff_\mathrm{mem}$ is the membrane force that results from the
Helfrich energy with bending modulus $k_b$, and a tension term that acts
as a Lagrange multiplier to satisfy the local inextensibility condition.
Following~\cite{vee-gue-zor-bir2009}, we modify the bending tension and
forces as
$\ff_\mathrm{mem} = \ff_\mathrm{ben} +
\ff_\mathrm{ten}$, where
\begin{align}
  \label{eqn:membraneForces}
  \ff_\mathrm{ben} = -k_b \xx_{ssss}, \qquad
  \ff_\mathrm{ten} = (\sigma \xx_s)_s.
\end{align}
The physical tension is $\Lambda = \sigma + 1.5 \kappa^2$, where
$\kappa$ is the membrane curvature.

Semi-permeability is introduced by modifying the vesicle velocity from
the standard choice. Instead of using the no-slip boundary condition, we
assume the fluid flux through the boundary is proportional to the
mechanical force~\cite{yao-mor2017}. That is, 
\begin{align}
  \label{eqn:vesicleFlux}
  \uu - \dot{\xx} = - \beta (\ff_\mathrm{mem} \cdot \nn) \nn, \qquad
  \xx \in \gamma,
\end{align}
where $\beta$ is the semi-permeability rate.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{Integral equation formulation}
Since the fluid satisfies the Stokes equations, it can be represented as
layer potential. Given the interfacial boundary
conditions~\eqref{eqn:jump}, the fluid velocity is
\begin{align}
  \uu(\xx) = \uu_\infty(\xx) + \SS[\ff_\mathrm{mem}](\xx), \quad
    \xx \in \Omega,
\end{align}
where $\SS$ is the single-layer potential
\begin{align}
  \SS[\ff](\xx) = \frac{1}{4\pi\mu} \int_{\gamma} \left(
    -\mathbf{I} \log\rho + \frac{\rr \otimes \rr}{\rho^2} \right)
    \ff(\yy) ds_{\yy},
\end{align}
with $\rr = \xx - \yy$ and $\rho = \norm{\rr}$. Imposing the flux
condition~\eqref{eqn:vesicleFlux} and the inextensibility condition, the
vesicle velocity satisfies
\begin{align}
  \label{eqn:vesVelocity}
  \dot{\xx} &= \uu_\infty(\xx) + \beta (\ff_\mathrm{mem}\cdot\nn)\nn
  + \SS [\ff_\mathrm{mem}](\xx),  \quad
  \xx_s \cdot \dot{\xx}_s = 0.
\end{align}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{Dimensionless equations}
The inextensibility condition guarantees that the vesicle's length, $L$,
is conserved, while its area, $A$, can change due to semi-permeability.
Therefore, the reduced area $\nu = 4\pi A/L^2 \in (0,1]$ is not
conserved. We non-dimensionalize~\eqref{eqn:vesVelocity} by introducing
the length scale $R_0 = L/2\pi$ and defining the time scale $\tau = \mu
R_0^3/k_b$, the tension scale $Q = k_b/R_0^2$, and the permeability
scale $W = R_0/\mu$. The background velocity $\uu_\infty$ is
non-dimensionalized with $R_0/\tau$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{Area dynamics}
Following the analysis of Veerapaneni et al.~\cite{vee-raj-bir-pur2009},
we compute the steady state shape of a semi-permeable vesicle in a
quiescent flow. The steady state shape is a stationary point of the
Lagrangian
\begin{align}
  \mathcal{L} = \frac{1}{2}\int_{\gamma} \kappa^2 \, ds +
    \Lambda \left(\int_{\gamma} ds  - L \right),
\end{align}
where the Lagrange multiplier $\Lambda$ is the tension. Note the absence
of the pressure Lagrange multiplier since the vesicle's area fluctuates.
Taking the variation of $\mathcal{L}$ with respect to $\gamma$ and
setting it to zero, the steady state shape satisfies $\kappa_{ss} +
\frac{1}{2}\kappa^3 - \Lambda \kappa = 0$, with initial conditions
$\kappa(0) = \kappa_0$ and $\kappa_s(0) = 0$.  Integrating once, we have
\begin{align}
  \frac{\kappa_s^2}{2} + \frac{\kappa^4}{8} - 
    \frac{\Lambda}{2}\kappa^2 = \frac{\kappa_0^4}{8} - 
    \frac{\Lambda}{2}\kappa_0^2.
  \label{eqn:curveODE}
\end{align}
Assuming $\kappa_s \neq 0$,~\eqref{eqn:curveODE} is separable, and can
be solved analytically in terms of the {\em EllipticF} function using
Mathematica. Unlike in the impermeable case~\cite{vee-raj-bir-pur2009},
the solution of~\eqref{eqn:curveODE} is not periodic unless $\kappa(s) =
\kappa_0$. Therefore, the steady state shape of a semi-permeable vesicle
in a quiescent flow is a circle with radius $L/2\pi$. Therefore, the
steady state curvature and tension are
\begin{align}
  \kappa = \frac{2\pi}{L}, \: \Lambda = \frac{2\pi^2}{L^2}.
  \label{eqn:SSshape}
\end{align}
Note that in contrast to a impermeable vesicle, the tension is positive
rather than negative.

We also compute the transient area of the vesicle. Since the area of the
vesicle is
\begin{align}
  A(t) = \frac{1}{2}\int_{\gamma} (\xx \cdot \nn)\, ds,
  \label{eqn:area}
\end{align}
its derivative includes terms due to the time derivative of $\xx \cdot
\nn$ and a term due to interface stretching~\cite{lai-tse-hua2008}, but
this latter term term is zero because of the local inextensibility
condition. Therefore,
\begin{align}
  \dot{A}(t) =
  \frac{1}{2} \int_{\gamma} (\dot{\xx} \cdot \nn)\, ds  + 
  \frac{1}{2} \int_{\gamma} (\xx \cdot \dot{\nn})\, ds.
\end{align}
Applying integration by parts, the two integrals agree, and we have
%\begin{align}
%  \int_\gamma (\xx \cdot \dot{\nn})\, ds 
%    = \int_\gamma \left(\xx \cdot \dot{\xx}_s^\perp \right)
%    = -\int_\gamma \left(\xx^\perp \cdot \dot{\xx}_s\right)\, ds 
%    = \int_\gamma \left(\xx_s^\perp \cdot \dot{\xx}\right)\, ds 
%    = \int_\gamma (\dot{\xx} \cdot \nn)\, ds.
%\end{align}
\begin{align}
  \dot{A}(t) = \int_\gamma \left(\beta (\ff \cdot \nn)\nn 
    + \SS[\ff]\right) \cdot \nn\, ds 
  = \beta \int_\gamma (\ff \cdot \nn)\, ds,
\end{align}
where we have used the incompressibility of the single-layer potential.
Repeatedly applying integration by parts, we have
\begin{align}
  \dot{A}(t) &= \beta \int_\gamma \left( -\xx_{ssss} + 
    (\sigma \xx_s)_s \right) \cdot \nn \, ds \\
  &= -\beta \int_\gamma \left(-\xx_{sss} + \sigma \xx_s 
    \right) \cdot \kappa \xx_s \, ds \\
  &= -\beta \int_\gamma \left(\xx_{ss} \cdot 
    (\kappa \xx_s)_s + \kappa \sigma \right) \, ds \\
  &= -\beta \int_\gamma \left(\kappa (\xx_{ss} \cdot \xx_{ss}) + 
    \kappa_s (\xx_{ss} \cdot \xx_s) + \kappa \sigma \right) 
    \, ds \\
  &= -\beta \int_\gamma \left(\kappa^3 + \kappa \sigma \right) 
    \, ds = \beta \int_\gamma \left(
    \frac{\kappa^3}{2} - \kappa \Lambda \right) \, ds.
  \label{eqn:areaROC}
\end{align}
Note that~\eqref{eqn:SSshape} and are consistent since the constant
curvature and tension values cause the integrand to vanish.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{Discretization in Space}
By using a boundary integral equation formulation, the unknowns are the
vesicle's position and tension that we track at $N$ discretization
points. Derivatives are computed with Fourier differentiation, and the
eighth-order quadrature~\cite{alp1999} is applied to the single-layer
potential. For constricted flows, nearly-singular integrals are computed
with an interpolation-based quadrature rule~\cite{qua-bir2014}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{Discretization in Time}
To eliminate stringent time step restrictions, we modify a time stepping
method that discretizes high-order derivatives
semi-implicitly~\cite{vee-gue-zor-bir2009}. We introduce the notation
\begin{alignat}{3}
  &B[\xx]\ff = -\frac{d^4}{ds^4} \ff,  \qquad
  &&T[\xx]\sigma = (\sigma \xx_s)_s, \\
  &D[\xx]\ff = \xx_s \cdot \ff_s, 
  &&P[\xx]\ff = (\ff \cdot \nn) \nn,
\end{alignat}
for the differential and projection operators. Letting $B^N$ be the
bending operator due to the vesicle configuration at time $t^N$, and
similar notation for the other operators, we use the first-order
semi-implicit time stepping method
\begin{alignat}{3}  
  \frac{\xx^{N+1} - \xx^N}{\Delta t} &= \uu_\infty(\xx^N) 
  &&+ \beta P^N\left(B^N\xx^{N+1} + T^N\sigma^{N+1}\right) \nonumber \\
  & &&+ \SS^N\left(B^N\xx^{N+1} + T^N\sigma^{N+1}\right), 
\end{alignat}
and  the inextensibility condition is discretized as
\begin{align}
  D^N\xx^{N+1} &= 1.
\end{align}
To achieve second-order accuracy over long time horizons, we modify an
adaptive spectral deferred correction time stepping
method~\cite{qua-bir2016}. Since the area of the vesicle is not
conserved, we only use the length of the vesicle to estimate the error
which determines if a time step is accepted or not and determines the
subsequent time step size.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Examples}
We consider a single semi-permeable vesicle in quiescent, shear, and
parabolic flows. We also consider a semi-permeable vesicle passing
through a constriction. For each example, all vesicles have a length
$L$, but the area is not conserved because of semi-permeability. We vary
the initial vesicle reduced area, the permeability rate, and the flow
rate. For all simulations, the bending modulus is $10^{-19}J$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{Quiescent Flow} 
We consider a semi-permeable elliptical vesicle with initial reduced
area $\nu = 0.65$ and several different permeability rates in a
quiescent flow. Based on our analysis, the steady state shape is a
circle with constant tension $\Lambda = 2\pi^2/L^2$.
By~\eqref{eqn:areaROC}, the area change is proportional to $\beta$, and
we expect both the time for the vesicle to deviate from its initial
area, and the time to reach its steady state to be proportional to
$\beta^{-1}$. In Table~\ref{tbl:ellipseRelaxTimes}, we report the time,
$t_\mathrm{steady}$, required for the vesicle's area to be within 1.3\%
of its maximum area, and the time, $t_\mathrm{inflate}$, required for
the vesicle's area to exceed 1.3\% of its initial area. Both times are
proportional to $\beta^{-1}$. We plot the area of a vesicle with six
different permeability rates in Figure~\ref{fig:ellipseArea}. The area
is computed directly from the vesicle shape and the time axis is scaled
by $\beta$. The plot verifies that the time for the vesicle to reach a
circular shape is proportional to $\beta^{-1}$. We also
solve~\eqref{eqn:areaROC} for the $\beta=10$ simulation and superimpose
its solution with dots on Figure~\ref{fig:ellipseArea}.
\begin{table}[htp]
  \centering
  \caption{\label{tbl:ellipseRelaxTimes} The time for a semi-permeable
  vesicle to its steady state ($t_\mathrm{steady}$) and to begin
  inflating ($t_\mathrm{inflate}$).}
  \begin{tabular}{C{0.8cm}C{2cm}C{0.8cm}C{2cm}C{0.8cm}}
    $\beta$ & $t_\mathrm{steady}$ & ratio & $t_\mathrm{inflate}$ & ratio \\
    \midrule
    $10^{0}$  & $6.45 \times 10^{0}$ & ---  
              & $2.17 \times 10^{-2}$ & --- \\ 
    $10^{-1}$ & $2.00 \times 10^{1}$ & 3.10 & 
                $1.25 \times 10^{-1}$ & 5.76 \\
    $10^{-2}$ & $9.61 \times 10^{1}$ & 4.81 & 
                $1.59 \times 10^{0}$ & 12.72 \\
    $10^{-3}$ & $8.36 \times 10^{2}$ & 8.71 & 
                $2.19 \times 10^{1}$ & 13.77 \\
    $10^{-4}$ & $8.15 \times 10^{3}$ & 9.74 & 
                $2.14 \times 10^{2}$ & 9.77 \\
    $10^{-5}$ & $8.03 \times 10^{4}$ & 9.85 & 
                $2.20 \times 10^{3}$ & 10.14 \\
    \bottomrule
  \end{tabular}
\end{table}

\begin{figure}[htp]
  \centering
%  \ifTikz
  \input{figures/ellipseArea.tikz}
%  \fi
%  \includegraphics[width=0.8\linewidth]{figures/ellipseArea}
  \caption{\label{fig:ellipseArea} The area of an elliptical
  semi-permeable vesicle, with permeability rate $\beta$ in a quiescent
  flow. The time axis is scaled by the permeability rate $\beta$. Since
  the curves collapse to a single curve, the permeability rate $\beta$
  sets a relaxation time scale.}
\end{figure}

We further analyze the transient dynamics by consider the vesicle
velocity~\eqref{eqn:vesVelocity} for small $\beta$. The vesicle velocity
depends on two terms: (i) the permeability velocity, $\beta(\ff \cdot
\nn)\nn$, and (ii) the force balance velocity, $\SS[\ff]$. At early
times, since $\beta \ll 1$, the permeability velocity is much smaller
than the force balance velocity and the vesicle dynamics resembles that
of a relaxing impermeable vesicle. As the force balance velocity
decreases, the permeability velocity becomes significant, and the
vesicle begins to inflate to a circle. This results in the membrane
force converging to zero \todo[inline]{Is this right?}, as does the
velocity due to permeability and force balance. Letting $\beta =
10^{-3}$, the force balance velocity (blue) and the permeability
velocity (red) are plotted at three times in
Figure~\ref{fig:vesVelocity}. We also plot the velocity of an
identically initialized impermeable vesicle (green). At $t=0.5$, the
permeability velocity is small relative to the force balance velocity,
and the force balance velocity matches with the impermeable vesicle
velocity. At time $t=5$, all three velocities are comparable in size,
and the semi-permeable vesicle begins to inflate. Then, at time $t=50$,
the impermeable vesicle is at its steady state, while the semi-permeable
vesicle is still inflating. The bottom right plot of
Figure~\ref{fig:vesVelocity} is the maximum norm of the different
velocity contributions as a function of time, and the black vertical
lines correspond to the times reported in the other three subplots.

%  \input{figures/ellipseVelocity1.tikz}
%  \input{figures/ellipseVelocity2.tikz}
%  \input{figures/ellipseVelocity3.tikz}
%  \input{figures/ellipseVelocityNorms.tikz}
\begin{figure}[htp]
  \centering
%  \input{figures/ellipseVelocity1.tikz}
%  \input{figures/ellipseVelocity2.tikz}
%  \input{figures/ellipseVelocity3.tikz}
%  \input{figures/ellipseVelocityNorms.tikz}
  \includegraphics[width=0.48\linewidth]{figures/ellipseVelocity1.pdf}
  \includegraphics[width=0.48\linewidth]{figures/ellipseVelocity2.pdf}
  \includegraphics[width=0.48\linewidth]{figures/ellipseVelocity3.pdf}
  \includegraphics[width=0.48\linewidth]{figures/ellipseVelocityNorms.pdf}
  \caption{\label{fig:vesVelocity} The velocity contributions due to
  force balance (blue) and permeability (red). The green dashed line is
  the velocity of an impermeable vesicle with the same initial shape as
  the semi-permeable vesicle. The bottom right plot is the max norm of
  each contributor to the vesicle velocities. The three black lines
  correspond to the times reported in the other three plots.}
\end{figure}
